%%%%%%%%%%%%%%%% beginning of timing analysis of single cycle datapath


\begin{frame}[fragile]
\STitle{Timing Instructions on Single Cycle Machines}
\begin{tcolorbox}[enhanced,attach boxed title to top center={yshift=-3mm,yshifttext=-1mm},
  colback=red!5!white,colframe=red!75!black,colbacktitle=red!80!black,
  title=Try this,fonttitle=\bfseries,
  boxed title style={size=small,colframe=red!50!black} ]
Suppose memory units take 200 ps (picoseconds), ALUs 200 ps,
register files 100 ps, no delay on other units.

How long to execute the instructions on the Single Cycle Datapath?

  \end{tcolorbox}
\begin{itemize}

\item Branch take 400 ps, CBZ take 500 ps, R-format instructions 600 ps,
STUR 700 ps, LDUR 800 ps.
\item Clock period must be increased to 800 ps or more
\item Even worse when floating-point instructions are implemented
\item Idea: use a multicycle implementation
% and R format
\end{itemize}
\BNotes\ifnum\Notes=1
\begin{itemize}
\item Ask the students why the different types of instructions take the
  times listed, or work it out on the board.
\item See Figure 4.25 of the text.  The figure doesn't list branch, but
  it's the same entry as CBZ except it doesn't use the register file.
  \end{itemize}
\fi\ENotes
\end{frame}

\input 04-Single-Cycle-Processor-Implementation/single-cycle-timing-f22


\begin{frame}[fragile]
\STitle{Conclusion}
 \underline{\textbf{Lecture Summary}}
 \begin{itemize}
 \item Closely analyzing instruction execution times on Single Cycle Datapath 
\end{itemize}
 % \end{itemize}
 \underline{\textbf{Assigned Textbook Readings}}
\begin{itemize}
     \item \textbf{Read} 
     \begin{itemize}
     \item Chapter 4, Section 4.4
   % \item   Appendix C, Section C.2\footnote{caution: the control unit is for MIPS datapath not ARM datapath!}
     \end{itemize}
     \end{itemize}
    \underline{\textbf{Next Steps}}
    \begin{itemize}
     \item \textbf{Review} logic behind the control unit implementation. 
% \begin{itemize}
%     % \item Start thinking about the exercises in A3
%     % \item A3 will be released in two parts, first set of exercises focus on Sequential Circuits and FSM, second set of exercises on Data Representation
% \end{itemize}
\item \textbf{Attempt} questions in next week's tutorial. 
    \item \textbf{Ask} questions in office hours or the next tutorial.
 \end{itemize}

\end{frame}